What does the crowd believe?

### 2016-08-08 presentations/01_CogSci_2016/slides.Rmd
presentation shortened and polished

<span style = "color:firebrick">recap</span>

- 3 types of Bayesianism:
    - Bayesian versions of classical statistical tests
    - Bayesian modeling of data-generating process
    - cognitive models that assume that people "are Bayesian"

<span style = "color:firebrick">today</span>

- subjective beliefs for models of "Bayesian cognition"

## what, where from, where to?

<span style = "color:firebrick">the "priors consortium"</span>

<div style = "float:left; width:45%;">

  - Erin D. Bennett
  - Fabian Dablander
  - Judith Degen
  - Michael Franke
  - Noah Goodman

<div style = "float:right; width:45%;">

  - Justine Kao
  - Anthea Schoeller
  - Michael Henry Tessler
  - Ciyang Qing
  - ... ??? ...

<p>

<div>
<span style = "color:firebrick">disclaimer</span>

- this is mess in progress
- all follies are mine

# background

## "Bayesian cognition"

<span style = "color:firebrick">Bayes for learning</span>

- prior beliefs on what "structure" can be expected
    - (possibly innate) learning biases
- Bayes rule to learn from sparse input
    - language acquisition 
    - conceptualization, generalization

<span style = "color:firebrick">Bayes for performance</span>

- model task behavior as (partly) Bayesian inference
    - rationalistic explanation, not necessarily mechanism/process
- subject's beliefs about everyday events (task items) plays a role
    - reasoning from premisses, future predictions, etc.
    - language use and interpretation
        - vagueness, non-literal interpretation etc.

## example: vague <span style = "font-style: italic">many</span>

@@

@@

Joe eats more bugers that we would expect of him.

<span style = "color:firebrick">theory</span>

- look at expectation $p$ about burger consumption for the "kind of guy" that Joe is
    - $p(n)$ is the probability that a guy like Joe eats $n$ burgers
    - let $P$ be the cumulative distribution of $p$
- let $b$ be the number of burgers that Joe actually eats
- "Joe eats many burgers" is true iff $P(b)$ is bigger than a fixed threshold
    - that threshold is the same for all sentences of this kind
    - contextual vagueness from uncertainty about $p$

@@

## enter subjective beliefs

<span style = "color:firebrick">desideratum</span>

- compact, versatile respresentation of subjective beliefs
- non-expert beliefs about possibly mundane affairs
- cheap, fast and reliable ways of assessment

<span style = "color:firebrick">previous literature</span>

- take real-world frequencies as stand-in for subjective beliefs
- measure beliefs experimentally, e.g.:
    - infer parameters of Gaussian from a judgements of cumulative probability (Manski, 2004)
    - infer parameters and type of distribution from predictions (Tauber & Steyvers 2013)
    - binned slider ratings (Kao et al., 2014)

## binned slider ratings

@@

- discretize domain into bins; let subjects rate each bin
- take relative measure: normalize slider ratings per participant
- average (normalized) slider rating to reflect population-level belief
    - "wisdom of the crowds"

@@

<span style = "color:firebrick">issues</span>

- wanted: reliable & practical way of measuring beliefs
    - consistency across measures
- relation subjects' answers, their beliefs and the population-level aggregate?
    - what of between-subject variance in task data?

@@

<span style = "color:firebrick">approach</span>

- belief elicitation with different tasks (within-subject)
- Bayesian hierarchical model to infer latent subjective & "population-level beliefs"
    - "population-level belief" => central tendency of subjective beliefs
    - think: "mean of a Gaussian hyperprior"

# experiments

@@

- 20 participants recruited via MTurk
- each saw every condition of every task
- 8 items (from previous research)
- 3 task types:
    - slider ratings
    - number estimates
    - binary comparison (a.k.a. "lightning round")

@@

@@

<div align = 'center'>
  <img src="pics/priors_sliders.png" alt="priorsslider" style="width: 600px;"/>
</div>

@@

<div align = 'center'>
  <img src="pics/priors_numbers.png" alt="priorsnumbers" style="width: 600px;"/>
</div>

## binary choices

@@

<div align = 'center'>
  <img src="pics/priors_lightning.png" alt="priorslightning" style="width: 600px;"/>
</div>

@@

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_slider.png" alt="dataslider" style="width: 600px;"/>
</div>

@@

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_number.png" alt="dataslider" style="width: 600px;"/>
</div>

@@

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_choice.png" alt="dataslider" style="width: 600px;"/>
</div>

# model

@@

- is there a consistent population-level average that explains participants choices?
- is the average slider rating a good approximation of a latent population-level average?
- how much individual variance is there?

@@

<span style = "color:firebrick">specific (technical)</span>

- hierarchical group-level prior for subjective beliefs?
- link functions for task types?

## slider data

$s_{ijk} \in [0;1]$ - normalized slider ratings of subject $i$ for item $j$ and bin $k$

```{r, echo = FALSE}
source('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/process_data.R')
# bin_dat <- 
# choice_dat <- read.csv('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/choice_dat.csv')
# number_dat <- read.csv('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/number_dat.csv')
# 
# 
# ## slider rating data
# y.slider = array(0, dim = c(20,8,15))
# for (i in 1:nrow(bin_dat)){
#   y.slider[bin_dat$workerid[i] + 1, 
#            as.numeric(bin_dat$tag)[i],
#            bin_dat$bin_num[i]] = bin_dat$nresponse[i]
# }
# # y.slider = logit(add.margin(y.slider))
# 
# # number choice data
# y.number = array(0, dim = c(20,8,1))
# for (i in 1:nrow(number_dat)){
#   y.number[number_dat$workerid[i] + 1, 
#            as.numeric(number_dat$tag)[i],
#            1] = number_dat$chosen_bin[i]
# }
# y.number = y.number[,,1]
# 
# ## lightning round choice data
# choice_dat = choice_dat %>% mutate(condition = factor(ifelse(chosen_bin > unchosen_bin, unchosen_bin, chosen_bin)))
# choice_dat$condition = as.numeric(choice_dat$condition)
# y.choice = array(0, dim = c(20,8,5))
# higher   = array(0, dim = c(20,8,5))
# lower    = array(0, dim = c(20,8,5))
# for (i in 1:nrow(choice_dat)){
#   y.choice[choice_dat$workerid[i] + 1, 
#            as.numeric(choice_dat$tag)[i],
#            choice_dat$condition[i]] = ifelse(choice_dat$higher_chosen[i] , 1, 0)
# }
```

```{r}
dim(y.slider)
y.slider[1,1:4,1:5]
```

## number data

$n_{ij} \in \{1, \dots, 15\}$ - number of bin for number choice of subject $i$ and item $j$

```{r}
y.number[1:12,,]
```

## lightning choice data

$c_{ijl} \in \{ 0, 1\}$ - whether subject $i$ chose the higher bin in lightning round $l \in \{ 1, \dots, 5\}$ for item $j$

```{r}
y.choice[1,,]
```

@@

- what's more likely: John scored 10-12, or John scored 26-28 points?
- subjective belief says: most likely John scored 14-16
- lower bin is closer, so

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
</div>

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priors.png" alt="postPriors" style="width: 580px;"/>
</div>

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priorsSubj.png" alt="postSubjPriors" style="width: 580px;"/>
</div>

## impressions

<span style = "color:firebrick">from visual inspection</span>

- subjective beliefs differ from population-level mean (good!)
- avrgd normlzd slider ratings nicely approximate mean $Q_j$ (excellent!)

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_slider.png" alt="ppcSlider" style="width: 560px;"/>
</div>

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_number.png" alt="ppcNumber" style="width: 560px;"/>
</div>

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_choice.png" alt="ppcChoice" style="width: 560px;"/>
</div>

>>>>

# motivation

## point of departure

<span style = "color:firebrick">relevance</span>

- belief + intention + rationality $\Rightarrow$ action

<span style = "color:white"> &nbsp; </span>

<span style = "color:firebrick">problem</span>

- beliefs are latent, only act is directly observable

<span style = "color:white"> &nbsp; </span>
<span style = "color:white"> &nbsp; </span>
<span style = "color:white"> &nbsp; </span>

<span style = "color:firebrick">INSERT A COOL PICTURE??</span>

## mundane beliefs may matter

@@

<span style = "color:white"> &nbsp; </span>

@@

Joe eats more burgers that we would expect of him.

<span style = "color:white"> &nbsp; </span>
<span style = "color:white"> &nbsp; </span>
<span style = "color:white"> &nbsp; </span>

<span style = "color:firebrick">INSERT A COOL PICTURE</span>

@@

## how to determine beliefs?

<span style = "color:firebrick">real-world frequencies</span>

- may not exist or may not be known

<span style = "color:white"> &nbsp; </span>

<span style = "color:firebrick">experimental measures</span>

- give-number task & inference of parameterized distribution 
    - e.g., Manski (2004); Tauber & Steyvers (2013)
    - <span style = "font-style: italic">scoring rules</span>
- binned histogram slider rating task (Kao et al., 2014)

## binned histogram slider ratings

@@

take average normalized slider ratings to reflect population-level belief

@@

<span style = "color:firebrick">goal: scrutinize BH task</span>

- relate subjects' answers, their beliefs & population-level aggregate

@@

<span style = "color:firebrick">approach: hieararchical Bayesian modeling</span>

- data from multiple task types (within-subject)
- infer latent subjective & "population-level beliefs"
    - "population-level belief" => central tendency of subjective beliefs
    - think: "mean of a Gaussian hyperprior"

# experiment

@@

- 50 participants recruited via MTurk
- each saw every condition of every task
- 8 items (from previous research)
- 3 task types:
    - <span style = "color:firebrick">BH</span>: binned histogram
    - <span style = "color:firebrick">GAN</span>: give-a-number
    - <span style = "color:firebrick">PC</span>: paired comparison

@@

## BH task

@@

<div align = 'center'>
  <img src="pics/priors_sliders.png" alt="priorsslider" style="width: 750px;"/>
</div>

## GAN task

@@

<div align = 'center'>
  <img src="pics/priors_numbers.png" alt="priorsnumbers" style="width: 750px;"/>
</div>

## PC task

@@

<div align = 'center'>
  <img src="pics/priors_lightning.png" alt="priorslightning" style="width: 750px;"/>
</div>

@@

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_slider.png" alt="dataslider" style="width: 750px;"/>
</div>

@@

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_number.png" alt="dataslider" style="width: 750px;"/>
</div>

@@

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/data_choice.png" alt="dataslider" style="width: 750px;"/>
</div>

@@

- do average BH ratings approximate latent population-level averages well?

<span style = "color:white"> &nbsp; </span>

<span style = "color:firebrick">specific</span>

- relation: choice behavior, subjective & population-level beliefs

```{r, echo = FALSE}
load('//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/data/processed_data.RData')
```

<!--

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
</div>
-->

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priors.png" alt="postPriors" style="width: 750px;"/>
</div>

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/pop_priorsSubj.png" alt="postSubjPriors" style="width: 750px;"/>
</div>

## upshot

<span style = "color:white"> &nbsp; </span>

- subjective beliefs differ from population-level mean (good!)
- avrgd normlzd slider ratings reasonably approximate mean $Q_j$ (excellent!)

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_slider.png" alt="ppcSlider" style="width: 750px;"/>
</div>

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_number.png" alt="ppcNumber" style="width: 750px;"/>
</div>

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/ppc_choice.png" alt="ppcChoice" style="width: 750px;"/>
</div>

@@

# modeling details

@@

- what's more likely: John scored 10-12, or John scored 26-28 points?
- subjective belief says: most likely John scored 14-16
- lower bin is closer, so it's more likely to be chosen
